Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem \thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}

### 2017-07-12 Response.tex

Additional detail has been incorporated which discusses the specific solver used in computing the solutions. 
The description does not specifically mention all of the possible options in \texttt{fsolve} but rather gives the key components used in our solution.
This approach seems to be consistent with the literature, e.g.~\cite{ozimek2010a,stuart2010}.
The modified portion is copied below.

\begin{correction}
In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
\end{correction}

\item
    \begin{itshape}
Did the authors experiment with the number of multiple shooting stages in their example problems?  There must be some sort of "sweet spot" in number of stages for the example problems attempted.  If I missed it, I apologize, but I didn't see it.
\end{itshape}

We did spend considerable time in determining the appropriate number of stages. 
One key issue was that our implementation in software depended on dividing the total number of integration steps by the number of interior stages.
For simplicity, this reduced to possible number of stages to an even number. 
Additionally, we discovered that adding many stages tended to degrade the solution and cause difficulties in covergence. 
The solver would tend to ensure a continuous trajectory, by satisfying the interior stages, at the expense of the terminal stage. 
Some additional detail has been provided that mentions our choice of the number of stages and has been copied below.

\begin{correction}
Based on experimentation, we use four stages in our multiple shooting method.
This provided the best performance and convergence stability while minimizing the difficulties in additional interior constraints.
\end{correction}

### 2017-07-12 Response.tex

The discussion of generating periodic orbits in the three body problem has been removed.
It replaced with a small introduction mentioning the differential correction process and a citation to a relevant resource. 
The revised paragraph is copied below.

\begin{correction}
The first objective is to design a transfer trajectory from a planar periodic orbit about the \( L_1\) Lagrange point to a bounded orbit in the vicinity of the Moon.
The target region is created by choosing an initial condition of \( x_0 = \begin{bmatrix}1.05 & 0 & 0 & 0.35 \end{bmatrix}^T \) with \( \mu = 0.0125 \).
The target set is propagated over a period of \( t = \num{20} \) in non-dimensional units which corresponds to approximately \num{1.5} years.
\Cref{fig:moon_orbit} shows that the target set remains in the vicinity of the Moon, or \( m_2\), in the rotating reference frame. 
This type of orbit would be useful for a variety of mission scenarios.
The bounded trajectories of the vehicles and constant line of sight to both the Moon and the Earth would allow for constant communication for future manned missions and potential habitats.
The initial set is a planar periodic orbit about \( L_1\), which is generated using the process of differential correction of a linear approximation~\cite{koon2011}.
\end{correction}

\item
    \begin{itshape}
"We define a maximum magnitude of the thrust as umax = 0.75 and assume we can point the thrust in any direction within the plane." And  "we again assume a upper bound on the thrust magnitude of umax = 0.75."  What was the reason behind this? Isn't this a pretty large value of dimensional acceleration for your problem of interest?  Even if it's chosen for demonstrative purposes, there should be some sort of explanation as to why.  In the context of what electric propulsion systems can deliver, it could be arguable that this value isn't really "low-thrust" but "finite-thrust".
\end{itshape}

The maximum thrust magnitude was chosen to be approximately \SI{1}{\newton} for an assumed \SI{500}{\kilo\gram} spacecraft. 
While larger than the NEXT thruster mentioned earlier, a cluster of similar engines could provide the required thrust (assuming adequate available power). 
Additional detail has been included in the manuscript describing this thrust limit in dimensional units. 
\begin{correction}
Next, we determine the reachability set with the addition of a low thrust control input.
We define a maximum magnitude of the thrust as \( u_{max} = 0.75 \) and assume we can point the thrust in any direction within the plane. 
This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
This acceleration limit is approximately \( u_{max} = \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
Assuming a fixed spacecraft mass of \SI{500}{\kilo\gram}, this model defines a maximum thrust of \SI{1}{\newton}.
Currently, the NASA NEXT xenon thruster is able to provide approximately \SI{0.25}{\newton} of thrust, and a cluster of such engines could be used to provide the desired thrust used in this work~\cite{schmidt2008}.
The trajectories are generated from a fixed initial state of \( \vecbf{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).
This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit. 
\end{correction}

### 2017-07-12 manuscript.tex
Add detail on choosing the number of shoooting stages
Reviewer 1 Comment 31
Additional detail on dimensional units of thrust for L1 reach
Reviewer 1 Comment 32 and 27

In~\cref{fig:multiple shooting}, we show a schematic representation of the multiple shooting procedure.
We split the optimal control horizon into equal length subsegments such that the length of each segment is \( \frac{k}{n} \), where \( k, n \) are the total number of steps and number of stages, respectively.
Similarly, we divide the state and costate trajectories into \( n \) equal segments. 
To ensure continuity, additional interior constraints are incorporated as
\begin{subequations}
\begin{align}\label{eq:interior_constraints}
        \vecbf{x}_1^{-} - \vecbf{x}_1^{+} &= 0 , \\ 
        \vecbf{\lambda}_1^{-} - \vecbf{\lambda}_1^{+} &= 0 , \\
        \vecbf{x}_2^{-} - \vecbf{x}_2^{+} &= 0 , \\ 
        \vecbf{\lambda}_2^{-} - \vecbf{\lambda}_2^{+} &= 0 , \\
        &\vdots \\
        \vecbf{x}_{n-1}^{-} - \vecbf{x}_{n-1}^{+} &= 0 , \\ 
        \vecbf{\lambda}_{n-1}^{-} - \vecbf{\lambda}_{n-1}^{+} &= 0.
\end{align}
\end{subequations}
Using the multiple shooting method reduces the sensitivity of the terminal states, \( \vecbf{x}_i^{+}, \vecbf{\lambda}_i^{+}\), to variations of the initial states, \(\vecbf{x}_{i-1}^{-}, \vecbf{\lambda}_{i-1}^{-}\).
As a result, the design vector \( \vecbf{\chi} \) is augmented with the additional interior initial conditions, \( \vecbf{x}_i^{-}, \vecbf{\lambda}^{-}\).
Similarly, the constraint vector is augmented with the additional interior constraints defined in~\cref{eq:interior_constraints}.
The multiple shooting algorithm now varies the design vector \( \vecbf{\chi}\) to ensure that the constraints in \( \vecbf{G}(\vecbf{\chi}) \) are satisfied.
In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
% tpbvp is solved using a shooting method (fsolve in matlab)
% determine a design vector X of variables we need to determine
% define a constraint vector G(X) that we need to satisfy
% multiple shooting has interior points of continuity constraints
% sensitivity of terminal state is used to correct the initial condition (cite a book/paper)
% plot of multiple shooting ( stages)
\section{Numerical Example}\label{sec:simulation}
We present two numerical simulations in the Earth-Moon system to demonstrate the transfer procedure.
These simulations enable the spacecraft to depart from the natural dynamics through the use of low-thrust propulsion.
The reachability set on the \Poincare section allows for a straightforward method of determining transfer opportunities.
The first example is a transfer from a periodic orbit of the \( L_1 \) Lagrange point to a fixed orbit of the moon.
This example uses a single iteration of the reachable set computation in the design of the transfer.
The second example is a transfer from geostationary orbit of the Earth to a period orbit of \( L_1 \). 
This examples demonstrates the ability to extend the reachability process to multiple iterations, to allow for a much larger and more general transfer.
With both examples it is possible to depart from the vicinity of the Earth to a Moon orbit via a series of reachable sets defined on \Poincare~sections.

@@

% reachable set transfer
Next, we determine the reachability set with the addition of a low thrust control input.
We define a maximum magnitude of the thrust as \( u_{max} = 0.75 \) and assume we can point the thrust in any direction within the plane. 
This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
The trajectories are generated from a fixed initial state of \( \vecbf{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).
This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit.

>>>>

In~\cref{fig:multiple shooting}, we show a schematic representation of the multiple shooting procedure.
We split the optimal control horizon into equal length subsegments such that the length of each segment is \( \frac{k}{n} \), where \( k, n \) are the total number of steps and number of stages, respectively.
Similarly, we divide the state and costate trajectories into \( n \) equal segments. 
To ensure continuity, additional interior constraints are incorporated as
\begin{subequations}
\begin{align}\label{eq:interior_constraints}
        \vecbf{x}_1^{-} - \vecbf{x}_1^{+} &= 0 , \\ 
        \vecbf{\lambda}_1^{-} - \vecbf{\lambda}_1^{+} &= 0 , \\
        \vecbf{x}_2^{-} - \vecbf{x}_2^{+} &= 0 , \\ 
        \vecbf{\lambda}_2^{-} - \vecbf{\lambda}_2^{+} &= 0 , \\
        &\vdots \\
        \vecbf{x}_{n-1}^{-} - \vecbf{x}_{n-1}^{+} &= 0 , \\ 
        \vecbf{\lambda}_{n-1}^{-} - \vecbf{\lambda}_{n-1}^{+} &= 0.
\end{align}
\end{subequations}
Using the multiple shooting method reduces the sensitivity of the terminal states, \( \vecbf{x}_i^{+}, \vecbf{\lambda}_i^{+}\), to variations of the initial states, \(\vecbf{x}_{i-1}^{-}, \vecbf{\lambda}_{i-1}^{-}\).
As a result, the design vector \( \vecbf{\chi} \) is augmented with the additional interior initial conditions, \( \vecbf{x}_i^{-}, \vecbf{\lambda}^{-}\).
Similarly, the constraint vector is augmented with the additional interior constraints defined in~\cref{eq:interior_constraints}.
Based on experimentation, we use four stages in our multiple shooting method.
This provided the best performance and convergence stability while minimizing the difficulties in additional interior constraints.
The multiple shooting algorithm now varies the design vector \( \vecbf{\chi}\) to ensure that the constraints in \( \vecbf{G}(\vecbf{\chi}) \) are satisfied.
In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
% tpbvp is solved using a shooting method (fsolve in matlab)
% determine a design vector X of variables we need to determine
% define a constraint vector G(X) that we need to satisfy
% multiple shooting has interior points of continuity constraints
% sensitivity of terminal state is used to correct the initial condition (cite a book/paper)
% plot of multiple shooting ( stages)
\section{Numerical Example}\label{sec:simulation}
We present two numerical simulations in the Earth-Moon system to demonstrate the transfer procedure.
These simulations enable the spacecraft to depart from the natural dynamics through the use of low-thrust propulsion.
The reachability set on the \Poincare section allows for a straightforward method of determining transfer opportunities.
The first example is a transfer from a periodic orbit of the \( L_1 \) Lagrange point to a fixed orbit of the moon.
This example uses a single iteration of the reachable set computation in the design of the transfer.
The second example is a transfer from geostationary orbit of the Earth to a period orbit of \( L_1 \). 
This examples demonstrates the ability to extend the reachability process to multiple iterations, to allow for a much larger and more general transfer.
With both examples it is possible to depart from the vicinity of the Earth to a Moon orbit via a series of reachable sets defined on \Poincare~sections.

@@

% reachable set transfer
Next, we determine the reachability set with the addition of a low thrust control input.
We define a maximum magnitude of the thrust as \( u_{max} = 0.75 \) and assume we can point the thrust in any direction within the plane. 
This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
This acceleration limit is approximately \( u_{max} = \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
Assuming a fixed spacecraft mass of \SI{500}{\kilo\gram}, this model defines a maximum thrust of \SI{1}{\newton}.
Currently, the NASA NEXT xenon thruster is able to provide approximately \SI{0.25}{\newton} of thrust, and a cluster of such engines could be used to provide the desired thrust used in this work~\cite{schmidt2008}.
The trajectories are generated from a fixed initial state of \( \vecbf{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).
This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit.
